---
title: "1a. Yield Data Processing"
subtitle: "Created January 12, 2023"
author: "PME"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
  pdf_document: default
  word_document: default
---

Data processing for the Brookings Alt-Rotation Yield Dataset.

Goals:

1. Make new factors and combine with rotation characteristics
2. Impute missing values and flag as such
3. Estimate protein and calorie production 

# Load
```{r message=FALSE}
renv::install('Matrix')  # for some reason this doesn't play well with renv. 
libs = c(
  'here',
  'magrittr',
  'tidyr',
  'lme4'
)

for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    library(i, character.only=TRUE)
  }
}

se = function(x) sd(x)/sqrt(length(x))
```

```{r}
in_dir = file.path('Data', 'Inputs')
in_df = 'Master.csv'
in_rot = "rotation_metrics.csv"

out_dir = 'Results'
out_data = 'Data'

df = here(in_dir, in_df) %>% 
  read.csv(stringsAsFactors=FALSE)

rot = here(in_dir, in_rot) %>% 
  read.csv(stringsAsFactors=FALSE)
```

## Set column names and make factors
```{r}
names(df) %<>% gsub('_KGHA', '', .)
is_factor = c('REP', 'CYCLE', 'ROTATION', 'YEAR', 'CURRENT', 'PREVIOUS')
df[is_factor] %<>% lapply(as.factor)

type_convert = function(x, as.is=TRUE) type.convert(x, as.is=as.is)
```

Make a factor, year in cycle
```{r}
df$CYCLE_YEAR = with(df, ((type_convert(YEAR)-1) %% 4) + 1) %>% 
  factor
```


### Investigate missing data
```{r}
table(df$CURRENT, is.na(df$YIELD)) %>% 
  cbind(PCT = round(100*(.[, 2]/rowSums(.)), 1))
```
Sunflower and canola had almost 50% failure. For canola, this was for artificial reasons - 
winter canola is poorly adapted to the region. So we will impute yields, but selectively ignore
those years depending on analysis.

For sunflower, yields in general were marred by artifacts of small plots - things like bird and weevil damage, so we will drop all sunflower-containing rotations. 

```{r}
missing_data = df[is.na(df$YIELD), c('CURRENT', 'YEAR')] %>% 
  table %T>%
  print
```
Canola had a total failure in 2002, 2007, 2008, and 2013-2016 - just under half
of years. This was due to poor overwintering - winter canola was grown which is 
not well adapted to the region, though spring canola is well-adapted. Therefore, 
yields are good when recorded.

Sunflower had a total failure in 2002, 2006, 2007, 2009, 2012, 2013, and 2016 - again, just under half of years. This was due to bird and weevil damage that make outsized
impacts on experimental plots. Yields when recorded are problematic, so will drop.

Canola and sunflower both failed in 2002, 2007, 2013, and 2016. 

Pea had a total failure in 2005. 

Flag years with missing data at crop and rotation level. 
```{r}
# number of plots with each crop in a given year
crop_plots = subset(df, YEAR=='2001') %>% 
  with(tapply(CURRENT, CURRENT, length))

# identify years with missing crop data
missing_data %<>% 
  data.frame
missing_data$CROP_PLOTS = crop_plots[missing_data$CURRENT]
missing_data$TOTAL_FAIL = with(missing_data, CROP_PLOTS == Freq)

# identify rotation -years with any missing data
missing_data %<>% merge(df[, c('ROTATION', 'CURRENT')] %>% unique, 
                        by='CURRENT') %>% 
  .[, c('CURRENT', 'ROTATION', 'YEAR', 'TOTAL_FAIL')] %>% 
  pivot_wider(names_from='CURRENT', 
              values_from='TOTAL_FAIL')
missing_data$HAS_CROP_FAIL = apply(missing_data[, -c(1:2)], 1, function(x) {
  ifelse(is.na(any(x)), FALSE, any(x))
})
missing_data %<>% 
  .[c('ROTATION', 'YEAR', 'HAS_CROP_FAIL')] %>% 
  unique


df %<>% merge(missing_data,
              by=c('YEAR', 'ROTATION'))
```

# Impute missing crop yields
Impute missing data by borrowing information from other treatments via BLUP. This assumes a rotation*previous crop effect, a year effect, and a rep effect on each crop's yield and shrinks estimates to the mean yield of a crop depending on how much additional information is available. 

Flag these values, of course. 
```{r}
mm = lmer(YIELD ~ (CURRENT|ROTATION:PREVIOUS) + (CURRENT|CYCLE:CYCLE_YEAR) + (CURRENT|REP:PLOT), 
          data=df, 
          control=lmerControl(calc.derivs=FALSE))
imputed_yield = predict(mm, 
                        df, 
                        allow.new.levels=TRUE)

# flag missing values as imputed and add to dataframe
df$IMPUTED = is.na(df$YIELD)
df[df$IMPUTED, 'YIELD'] = imputed_yield[df$IMPUTED]
```


```{r}
n_imp = sum(df$IMPUTED)
pct_imp = {n_imp/nrow(df)} %T>%
  print
```
This imputed `r n_imp` values, which are `r round(pct_imp, 3)*100`% of observations.


Drop sunflower. 
```{r}
df %<>% .[!grepl('Sf', df$ROTATION), ] %>% 
  droplevels
```

# Export
```{r}
here(out_data,
     'crop performance.csv') %>% 
  write.csv(df, ., row.names=FALSE)
```


# Whole Rotation Metrics
```{r}
facts = 
  c('YEAR', 
    'ROTATION', 
    'REP', 
    'CYCLE',
    'CYCLE_YEAR',
    'NSPECIES',
    'CROP_DIV',
    'LENGTH',
    'AGE')
ff = paste(facts, collapse='+') %>% 
  paste('YIELD', ., sep='~') %>% 
  formula
yields = df %>% 
  aggregate(ff, data=., mean)
```

## Export
```{r}
here(out_data,
     'rotation performance.csv') %>% 
  write.csv(yields, ., row.names=FALSE)
```

