---
title: "Process Weather Data"
subtitle: "Created July, 2020"
author: "PME"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
  pdf_document: default
  word_document: default
---

```{r set_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, 
                      warning=FALSE,
					  fig.align='center')
```

# Overview

Weather data of the rotation experiment at NCARL/ESDSWRF 2000-2016.

Goals:

1. Mean temperature
  1. Flowering
  2. Intervals throughout the growing season
2. Precipitation
  1. At intervals throughout the growing season

Flowering Windows:

- Winter Wheat	June 2-8
- Pea	June 8-14
- Spring Wheat	June 10-16
- Oats	June 14-20
- Canola	June 16-22
- Soybean	July 2-8
- Corn	July 24-30

Precip categories - monthly, April 15-Sept 15

Mean Temp - monthly, April 15-Sept 15


# Load
## Libraries and Data
```{r}
libs = c('ggplot2', 
         'magrittr', 
         'here', 
         'lme4', 
         'tidyr', 
         'gridExtra',
         'lattice')

in_dir = file.path('Data', 'Inputs')
in_df = '2000-2016 Weather.csv'

out_dir = 'Data'
out_df = 'Weather Summary.csv'
```


```{r}
for (i in libs) {
  if (!(require(i, character.only=TRUE))) { # install if necessary
    install.packages(i)
    library(i, character.only=TRUE)
  }
}


df = here(in_dir, in_df) %>% 
  read.csv

str(df)
```

### Munge and summarize
Won't worry about leap years, as a day difference shouldn't be a big deal. 
```{r}
df$YEAR = strsplit(df$DATE, '/') %>% 
  sapply('[', 3)
df$DATE %<>% as.Date(format='%m/%d/%Y')

flower_intervals = rbind(
  Wwheat = c(152, 160),  # DOY
  Pea = c(158, 166),
  Swheat = c(160, 168),
  Oat = c(164, 172),
  Can = c(166, 174),
  Soy = c(181, 190),
  Corn = c(204, 212)
  # spring = c(162, 171),
  # soy = c(182, 191),
  # corn = c(202, 211)
) %>%
  set_colnames(c('start', 'end'))

summary_intervals = rbind(
  MarApr = c(75, 105),
  AprMay = c(106, 135),
  MayJun = c(136, 166),
  JunJul = c(167, 196),
  JulAug = c(197, 227),
  AugSep = c(228, 258)
) %>% 
  set_colnames(c('start', 'end'))

```

# Summarize
## Functions
```{r}
wide_to_tall = function(x, val='value') {
  x %<>% as.data.frame
  x$YEAR = rownames(x) %>% 
    as.character
  x %<>% pivot_longer(cols=where('is.numeric'),
                      names_to='INTERVAL',
                      values_to=val)
  return(x)
}

mat_to_df = function(x, val='value') {
  x %<>% as.data.frame
  colnames(x) %<>% paste(val, sep='_')
  return(x)
}
```

## Flowering temps
```{r}
df_annual = split(df, df$YEAR)
flower_temps = apply(flower_intervals, 1, function(x) {
  sapply(df_annual, function(y) {
    ss = y$DOY >= x['start'] & y$DOY <= x['end']
    mean(y[ss, 'MAX_T'], na.omit=TRUE)
  })
})
flower_temps %>% round(1)
```


```{r}
levelplot(t(cor(flower_temps)), at=seq(-1, 1, by=0.05),
                   main='correlation among max temp at flowering')
```

```{r}
flower_temps %<>% mat_to_df('FLOWER_AVG_MAXT')
```


## Mean temp at intervals
```{r}
interval_temps = apply(summary_intervals, 1, function(x) {
  sapply(df_annual, function(y) {
    ss = y$DOY >= x['start'] & y$DOY <= x['end']
    mean(y[ss, 'AVG_T'], na.omit=TRUE)
  })
})
interval_temps %>% round(1)
```

```{r}
levelplot(cor(interval_temps), at=seq(-.5, .5, by=0.05),
                   main='correlation among average temperature')
```

```{r}
interval_temps %<>% mat_to_df('MEANT')
```


## Total precip at intervals
```{r}
interval_precip = apply(summary_intervals, 1, function(x) {
  sapply(df_annual, function(y) {
    ss = y$DOY >= x['start'] & y$DOY <= x['end']
    sum(y[ss, 'PRECIP'], na.omit=TRUE)
  })
})
interval_precip %>% round(1)
```

```{r}
levelplot(t(cor(interval_precip)), at=seq(-.5, .5, by=0.05),
                   main='correlation among total precipitation')
```

```{r}
interval_precip %<>% mat_to_df('PRECIP')
```

# Combine and export
```{r}
weather_summary = cbind(flower_temps, interval_precip, interval_temps) %>% 
  data.frame(
    YEAR = rownames(.),
    .
  ) %>% 
  set_rownames(NULL)
```

## Export
```{r}
here(out_dir, out_df) %>% 
  write.csv(weather_summary, 
            .,
            row.names=FALSE)
```
