---
title: "Part 2: Productivity vs Stability"
subtitle: "Created January, 2023"
author: "PME"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
  pdf_document: default
  word_document: default
---
# Overview

Steps:

1. Load
2. Assess productivity of rotations
  1. Calculate rotation yield within rotation-block-year
  2. ANOVA
3. Environmental index analysis of rotations
  1. Calculate environmental index = weighted average yield across rotations within block-year.
  2. Yield vs environmental index
2. Assess expected yield (means)
3. Assess buffering (environmental index)
  1. Rank rep-years by means of all crops
  2. Run environmental index analysis
3. Assess productivity-stability Relationships

## Load
```{r message=FALSE}
libs = c(
  'here',
  'magrittr',
  'tidyr',
  'lme4',
  'car',
  'ggplot2',
  'gridExtra',
  'emmeans',
  'multcomp',
  'ggpmisc',
  'ggExtra',
  'ggtext'
)

for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    library(i, character.only=TRUE)
  }
}

se = function(x) sd(x)/sqrt(length(x))

# slightly tuned Paired2 from color brewer
rot_colors = c(
  CS     = '#444444', # dark gray
  CCaWwS = '#B2DF8A', # light green
  CSSwS  = '#33A02C', # green
  COWwS  = '#FB9A99', # light red
  CCSSw  = '#E31A1C', # red
  CPWwS  = '#a89ad0', # light purple
  CSSwP  = '#6951b6'  # purple
)

rot_colors = c('#E8601C',
               '#4EB265',
               '#1965B0')
rot_colors %<>% {c(
  CS = '#444444',
  CCaWwS = .[2],
  CSSwS = .[2],
  COWwS = .[3],
  CCSSw = .[3],
  CPWwS = .[1],
  CSSwP = .[1]
)}
```

```{r}
in_dir = 'Data'
in_crop = 'crop performance.csv'
in_rot = 'rotation performance.csv'

out_dir = 'Results'

crop = here(in_dir, in_crop) %>% 
  read.csv

rot = here(in_dir, in_rot) %>% 
  read.csv
```

ensure factors are flagged as such
```{r}
facts = 
  c('YEAR', 
    'ROTATION', 
    'REP',
    'CYCLE',
    'CYCLE_YEAR')
crop[, facts] %<>% lapply(as.factor)
rot[, facts] %<>% lapply(as.factor)

# rot_colors %<>% .[levels(crop$ROTATION)]
crop[, 'ROTATION'] %<>% factor(levels=names(rot_colors))
rot[, 'ROTATION'] %<>% factor(levels=names(rot_colors))

# Convert yield to Mg ha
crop$YIELD = crop$YIELD/1000
rot$YIELD = rot$YIELD/1000
```

# Rotation Performance
This is mean yield across crops within a rotation

## Model and diagnostics
```{r}
pred = 'ROTATION'
rand = '(1|CYCLE/CYCLE_YEAR) + (1|REP/CYCLE_YEAR)'
resp = 'YIELD'

mod = paste(resp, '~', pred, '+', rand) %>% 
  formula %>% 
  lmer(data=rot)
Anova(mod, test.statistic='F')
```
```{r}
qqPlot(residuals(mod))
```

```{r}
plot(mod)
```
```{r}
influencePlot(mod)
```
Looks good. 

## Results
```{r}
mod_emm = 
  emmeans(mod, pred) %>% 
  cld(p.adjust='fdr', Letters=letters, reverse=TRUE) %T>%
  print
```
```{r}
emmeans(mod, pred) %>% 
  pairs(adjust='fdr')
```


```{r}
group_letters = function(x) {
  grp = x$.group  %>%
    trimws %>% 
    strsplit(split="")
  maxgrp = unlist(x$.group) %>% 
    as.numeric %>% 
    max
  grp %<>% sapply(function(y) {
    y %<>% as.numeric
    y = maxgrp - y + 1
    y %<>% sort(decreasing=FALSE) %>% 
      letters[.] %>% 
      paste(collapse='')
    y
  })
  
  out = data.frame(
    x[1],
    GROUP=grp
  )
  return(out)
}

mean_se = function(...) {
  out = aggregate(..., function(x) {
    se = sd(x)/sqrt(length(x))
    mn = mean(x)
    c(
      mean=mn,
      se=se,
      se_high=mn + se,
      se_low=mn - se)
  }) |>
    lapply(unlist) |>
    data.frame()
  
  nn = length(names(out))
  nn = seq(nn-3, nn, by=1)
  names(out)[nn] = c('mean', 'se', 'err_high', 'err_low')
  return(out)
}
```


```{r}
productivity = mean_se(YIELD ~ ROTATION, data=crop) |>
  merge(mod_emm,
        by='ROTATION')

productivity_plt = 
  ggplot(productivity,
         aes(x=ROTATION,
             y=mean,
             fill=ROTATION)) +
  scale_fill_manual(values=rot_colors) +
  # scale_fill_brewer(palette='Paired') +
  geom_col() +
  geom_errorbar(aes(ymin=err_low,
                    ymax=err_high),
                width=0.1) +
  geom_text(aes(y=err_high+0.450,
                label=trimws(.group)),
            size=2.5) +
  labs(x=NULL,
       y='Average Yield<br>2001-2016 [Mg ha<sup>-1</sup>]',
       fill=NULL) +
  guides(fill='none') +
  theme(panel.background=element_blank(),
        axis.title=element_markdown())
productivity_plt
```
## Discussion
COWwS and CS were highest yielding. CCaWwS and CSSwS were lowest yielding.

# Stability

Via environmental index (mean yield within years, scaled across years)

## Calculate indices
```{r}
ei = aggregate(YIELD ~ ROTATION + REP + YEAR, data=crop, mean) %>% 
  aggregate(YIELD ~ REP + YEAR, data=., mean) # average annual block yields including imputed values
names(ei) %<>% gsub('YIELD', 'EI', .)
ei$EI %<>% scale

ei %<>% merge(rot, by=c('REP', 'YEAR'))
ggplot(ei, 
       aes(x=EI)) +
  geom_density()
```
## Model and diagnostics
```{r}
pred = 'EI*ROTATION'
rand = 'REP'
resp = 'YIELD'

mod = paste(resp, '~', pred, '+', rand) %>% 
  formula %>% 
  lm(data=ei)
mod_av = Anova(mod, test.statistic='F') %T>%
  print
```
A significant interaction and a significant main effect of rotation. 


```{r}
qqPlot(residuals(mod))
```

```{r}
plot(mod)
```
```{r}
influencePlot(mod)
```
Looks good. 

```{r}
ei_plt = 
  ggplot(ei,
         aes(x=EI,
             y=.data[[resp]],
             color=ROTATION)) +
  scale_color_manual(values=rot_colors) +
  geom_vline(aes(xintercept=0),
             color='gray',
             size=0.2) +
  geom_point(size=0.5) +
  geom_smooth(method='lm',
              formula='y~x',
              se=FALSE,
              size=0.3) +
  labs(x='Environmental Index [*σ*]',
       y='Average Yield<br>[Mg ha<sup>-1</sup>]',
       color=NULL) + 
  annotate(geom='text',
           x=-1.3,
           y=6.700,
           label=paste("italic(p)<0.001"),
           parse=TRUE,
           size=2.5) +
  annotate(geom='text',
           x=-1.3,
           y=6.100,
           label=paste("italic(r)^2==", round(summary(mod)$r.squared, 3)),
           parse=TRUE,
           size=2.5) +
  theme(panel.background=element_blank(),
        legend.key=element_blank(),
        axis.title=element_markdown())
ei_plt
```



```{r}
mod_emm = 
  emtrends(mod, 'ROTATION', 'EI') %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE) %T>%
  print
```
```{r}
pwpm(emtrends(mod, 'ROTATION', 'EI'), adjust='fdr')
```


```{r}
stability_plt = 
  ggplot(mod_emm,
         aes(x=ROTATION,
             y=EI.trend,
             fill=ROTATION)) +
  scale_fill_manual(values=rot_colors) +
  geom_col() +
  geom_errorbar(aes(ymin=lower.CL,
                    ymax=upper.CL),
                width=0.1) +
  geom_text(aes(label=trimws(.group),
                y=upper.CL+0.1),
            size=2.5) +
  labs(x=NULL,
       y='Weather Susceptibility<br>Index [a.u.]') +
  guides(fill='none') +
  theme(panel.background=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(vjust=10),
        text=element_text(size=10),
        axis.title=element_markdown())
stability_plt
```

# Productivity vs Stability
```{r}
pred = "EI:REP:ROTATION"
rep_ei = paste(resp, pred, sep='~') %>% 
  formula %>% 
  lm(data=ei) %>% 
  coef %>% 
  .[-c(1)] %>% 
  data.frame(do.call(rbind, strsplit(names(.), ':')),
             value=.) %>% 
  set_rownames(NULL) %>% 
  set_colnames(c('EI', 'REP', 'ROTATION', 'SLOPE'))
rep_ei$EI = NULL
rep_ei$REP %<>% gsub('REP', '', .) %>% 
  factor(levels=levels(ei$REP))
rep_ei$ROTATION %<>% gsub('ROTATION', '', .) %>% 
  factor(levels=levels(ei$ROTATION))

rep_prod = paste(resp, 'REP:ROTATION', sep='~') %>% 
  formula %>% 
  aggregate(ei, mean)

cv_prod = paste(resp, 'REP:ROTATION', sep='~') %>% 
  formula %>% 
  aggregate(ei, function(x) 100*sd(x)/mean(x)) %>% 
  set_colnames(c('REP', 'ROTATION', 'CV'))

prod_ei = merge(rep_prod, rep_ei, by=c('REP', 'ROTATION')) %>% 
  merge(cv_prod, by=c('REP', 'ROTATION'))

ps_mod = lm(SLOPE ~ YIELD, data=prod_ei)

ps_test = Anova(ps_mod) %T>%
  print
```

```{r}
quadrants = data.frame(
  x = c(3390, 4150, 3800, 3780)/1000,
  y = c(660, 660, 950, 380)/1000,
  label = c('Less\nProductive', 
            'More\nProductive',
            'Variable',
            'Stable')
)

convex_hull = split(prod_ei, prod_ei$ROTATION) %>% 
  lapply(function(x) {
    tt = x[, c('ROTATION', 'YIELD', 'SLOPE')] %>% 
      .[chull(.[, c(2, 3)]), ]
  }) %>% 
  do.call(rbind, .)

centers = sapply(c('YIELD', 'SLOPE'), 
                 function(x) tapply(prod_ei[, x], prod_ei$ROTATION, mean)) %>% 
  as.data.frame
centers$ROTATION = rownames(centers)
```

## Outliers
The estimate here is the percentage discrepancy of observed weather dependency
relative to expected weather dependency given a sample unit's productivity. 
```{r}
prod_ei$resid = residuals(ps_mod)
prod_ei$fit = fitted(ps_mod)

ei_ttest = with(prod_ei, tapply(resid/fit, ROTATION, t.test))
ei_ttest %<>% lapply(function(x) data.frame(diff = x$estimate,
                                           p_val = x$p.value)) %>% 
  do.call(rbind, .)
ei_ttest$p_val %<>% p.adjust(method='fdr')
ei_ttest[order(ei_ttest$diff),]
```
So COWwS is `r round(100*ei_ttest['COWwS', 'diff'], 1)`% less susceptible to 
weather than expected. 

```{r}
ei_ttest$SIG = ei_ttest$p_val < 0.05
```

## Plot

```{r fig.height=4, fig.width=4.5}
prod_ei$SIG = ei_ttest[prod_ei$ROTATION, 'SIG']
convex_hull$SIG = ei_ttest[convex_hull$ROTATION, 'SIG']

centers$LABEL = centers$ROTATION %>% 
  sapply(function(x) {
    if (ei_ttest[x, 'p_val'] <= 0.05) paste0(x, '*')
    else x
  })


ps_plt =
  ggplot(prod_ei, 
         aes(x=YIELD,
             y=SLOPE)) +
  # scale_color_brewer(palette='Paired') +
  scale_color_manual(values=rot_colors) +
  scale_fill_manual(values=rot_colors) +
  # scale_fill_manual(values=rep('black', nrow(centers))) +
  # scale_color_manual(values=rep('lightgray', nrow(centers))) +
  geom_vline(aes(xintercept=mean(YIELD)),
             color='gray',
             size=0.1,
             linetype='dashed') +
  geom_hline(aes(yintercept=mean(SLOPE)),
             color='gray',
             size=0.1,
             linetype='dashed') +
  geom_point(color='black',
             size=0.1,
             shape=20) +
  geom_smooth(method='lm',
              formula='y~x',
              alpha=0.1,
              color='black',
              size=0.3) +
  geom_polygon(data=convex_hull,
               aes(color=ROTATION,
                   fill=ROTATION),
               linewidth=0.2,
               # fill='darkgray',
               alpha=0.2) +
  geom_text(data=centers,
            aes(label=ROTATION),
            color='black',
            size=2) +
    annotate(geom='text',
           x=3.400,
           y=.900,
           label=paste("italic(p)<", 0.001),#round(ps_test['YIELD', 4], 3)),
           parse=TRUE,
           size=2.5) +
  annotate(geom='text',
           x=3.400,
           y=.930,
           label=paste("italic(r)^2==", round(summary(ps_mod)$r.squared, 3)),
           parse=TRUE,
           size=2.5) +
  geom_text(data=quadrants,
            aes(x=x,
                y=y,
                label=label),
            color='#666666',
            size=2.5) +
  guides(fill='none',
         color='none') +
    labs(x='Average Yield, 2001-2016 [Mg ha<sup>-1</sup>]',
         y='Weather Susceptibility Index [a.u.]',
         color=NULL) +
  theme(panel.background=element_blank(),
        legend.key=element_blank(),
        axis.title=element_markdown())
```


```{r fig.height=4, fig.width=4.5}
ps_plt
```


# Risk and Reward

In low, average, and high years, what is the yield of each rotation?

low = bottom 32%, defined by environmental index score (z-score) < -1.  
High = top 32%, defined by environmental index score (z-score) >= 1.  
```{r}
cutoff = qnorm(.75)
cutoff=1

gen_labels = function(cutoff) {
  q = pnorm(cutoff)
  
  tails = 1-q
  mid = 1-(2*q)
  
  perc = 100*c(tails, mid, tails) %>%
    round(2) %>% 
    abs
  
  paste(c('Worst', 'Middle', 'Best'),
        perc) %>% 
    paste0('%')
}
labs = gen_labels(cutoff)

ei$YEAR_TYPE = labs[2]
ei[ei$EI < -cutoff, 'YEAR_TYPE'] = labs[1]
ei[ei$EI >= cutoff, 'YEAR_TYPE'] = labs[3]

ei$YEAR_TYPE %<>% factor(levels=labs[3:1])

years_by_type = tapply(ei$YEAR, ei$YEAR_TYPE, function(x) length(unique(x)))

mod = lmer(YIELD ~ YEAR_TYPE * ROTATION + (1|REP/CYCLE_YEAR) + (1|CYCLE/CYCLE_YEAR), data=ei)

qqPlot(residuals(mod))
```


```{r}
Anova(mod, test.statistic='F')
```

```{r}
em = emmeans(mod, ~ ROTATION | YEAR_TYPE)

cld(em, Letters=letters, reversed=TRUE, adjust='fdr')
```

```{r}
tt = tapply(ei$YIELD, ei$ROTATION, mean) %>% 
  sort(decreasing=TRUE)

ei$ROTATION_ORDER = factor(ei$ROTATION, levels=names(tt))

errorbar_quantiles = function(x, range) {
  y = range/2
  med = quantile(x, 0.5)
  low = quantile(x, y)
  hi = quantile(x, 1-y)
  return(data.frame(
    y = med,
    ymin = low,
    ymax = hi
  ))
}


risk_plt = 
  ggplot(ei,
         aes(x=ROTATION_ORDER,
             y=YIELD,
             color=ROTATION)) +
  facet_wrap(~YEAR_TYPE, 
             ncol=1,
             strip.position='right') +
  scale_color_manual(values=rot_colors) +
  stat_summary(fun.data=errorbar_quantiles,
             geom='pointrange',
             fun.args=list(range=0.05),
             size=0.1) +
  stat_summary(fun.data=errorbar_quantiles,
               geom='pointrange',
               fun.args=list(range=0.5),
               linewidth=1.4,
               size=0.3) +
  # stat_summary(fun.data=errorbar_range,
  #              geom='pointrange') +
  labs(x=NULL,
       y='Yield [Mg ha<sup>-1]') +
  guides(color='none') +
  theme(panel.grid=element_blank(),
        panel.background=element_rect(fill='#f7f7f7'),
        axis.text.x=element_text(angle=90, 
                                 vjust=0.5, 
                                 hjust=1),
        axis.ticks=element_line(linewidth=0.1),
        strip.background=element_blank(),
        axis.title=element_markdown())
risk_plt
```
Emphasizing that COWwS yields are quite reliable in poor years, with low downside
risk. CS yields, meanwhile, can be excellent. 



# Export
```{r fig.height=5, fig.width=6.5}
p1 = productivity_plt +
  labs(tag='a)') +
  theme(axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=45, hjust=1, vjust=1.3),
        text=element_text(size=10))


p2 = ei_plt +
  guides(color='none') +
  labs(tag='b)') +
  theme(text=element_text(size=10))

p3 = stability_plt + 
  labs(tag='c)') +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        text=element_text(size=10))


p4 = ps_plt + 
  guides(color='none') +
  labs(tag='\nd)') +
  theme(text=element_text(size=12))


# grid.arrange(p1, p2, p3, p4, nrow=2)

marrangeGrob(grobs=list(p1, p2, p3, p4), 
             layout_matrix=rbind(c(1,1,4,4,4,4),
                                 c(2,2,4,4,4,4),
                                 c(3,3,4,4,4,4)),
             top=NULL)
```

```{r}
here('Results', 'Figure 2 - productivity and stability.jpg') %>% 
  jpeg(width=6.5, height=5, units='in', res=300)
marrangeGrob(grobs=list(p1, p2, p3, p4), 
             layout_matrix=rbind(c(1,1,4,4,4,4),
                                 c(2,2,4,4,4,4),
                                 c(3,3,4,4,4,4)),
             top=NULL)
dev.off()
```


```{r fig.height=4, fig.width=6.5}
p4 = ps_plt + 
  guides(color='none') +
  labs(tag='b)') +
  theme(text=element_text(size=8),
        axis.ticks=element_line(linewidth=0.1))

p5 = risk_plt +
  scale_y_continuous(breaks=seq(2, 8, by=2)) +
  labs(tag='a)') +
  theme(text=element_text(size=8))

marrangeGrob(grobs=list(p5, p4),
             layout_matrix=rbind(c(1,1,2,2,2,2),
                                 c(1,1,2,2,2,2),
                                 c(1,1,2,2,2,2)),
             top=NULL)
```

```{r}
here('Results', 'Figure 2 ALT - productivity and stability.jpg') %>% 
  jpeg(width=6.5, height=4, units='in', res=300)
marrangeGrob(grobs=list(p5, p4),
             layout_matrix=rbind(c(1,1,2,2,2,2),
                                 c(1,1,2,2,2,2),
                                 c(1,1,2,2,2,2)),
             top=NULL)
dev.off()
```

```{r}
here('Results', 'Figure 2 - EI vs Yield.csv') %>% 
  write.csv(prod_ei, ., row.names=FALSE)
```


