---
title: "Part 3: Overyielding"
subtitle: "Created March, 2023"
author: "PME"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding=encoding, output_dir=here::here('Results'))})
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: hide
  pdf_document: default
  word_document: default
---


# Overview

In Part 2, we saw that high-yielding rotations are often, but not always, less
reliable. Specifically, the CS rotation was high yielding but also quite
susceptible to weather variation. The COWwS rotation was as high-yielding as CS, 
but was as stable across weather-years as the most stable rotation, CCaWwS, and
the lowest-yielding rotation, CSSwS. 

What allows COWwS to be so high-yielding? Does corn and soybean yield better 
in this rotation than in CS?

Steps:

1. Load
2. Assess productivity of corn and soybean within rotation
  1. Test yield ~ rotation for cycles 2-4 - to allow legacy effects to accumulate.
  2. Contrasts/effects
    1. Corn - yield in CS + CCSSw vs other 4-year; CS vs COWwS
    2. Soy - yield in CS + CSSwS vs other 4-year; CS vs COWwS
    3. And both from diversity, richness, functional group
3. Also test Ww and Sw yield ~ rotation where existing. 
  


## Load
```{r message=FALSE}
libs = c(
  'here',
  'magrittr',
  'tidyr',
  'lme4',
  'car',
  'ggplot2',
  'gridExtra',
  'emmeans',
  'multcomp',
  'ggtext',
  'wesanderson'
)

for (i in libs) {
  if (!require(i, character.only=TRUE)) {
    renv::install(i)
    library(i, character.only=TRUE)
  }
}

se = function(x) sd(x)/sqrt(length(x))

# selected from color brewer Set1
crop_colors = c(
  Can =  "#4DAF4A",
  Corn = "#FF7F00",
  Oat = "#999999",
  Pea =  "#984EA3",  
  Soy =  "#377EB8",
  Swheat = "#F781BF",
  Wwheat = "#E41A1C"
)

rot_colors = c(
  CS     = '#4b4b4b', # dark gray
  CCaWwS = '#B2DF8A', # light green
  CSSwS  = '#33A02C', # green
  COWwS  = '#FB9A99', # light red
  CCSSw  = '#ae1415', # red
  CPWwS  = '#a89ad0', # light purple
  CSSwP  = '#6951b6'  # purple
)

rot_shapes = c(
  CCaWwS = 16, 
  CCSwS  = 17, 
  COWwS  = 15, 
  CPWwS  = 3, 
  CS     = 7, 
  CSSwP  = 8, 
  CSSwS  = 5
)

pseudo_palette_3 = c('#E8601C',
                     '#4EB265',
                     '#1965B0')

pretty_crop = c(
  Can = 'Canola',
  Corn = 'Corn',
  Wwheat = 'Winter\nWheat',
  Swheat = 'Spring\nWheat',
  Oat = 'Oat',
  Pea = 'Pea',
  Soy = 'Soybean'
)

```

```{r}
in_dir = 'Data'
in_crop = 'crop performance.csv'
in_rot = 'rotation performance.csv'

out_dir = 'Results'

df = here(in_dir, in_crop) %>% 
  read.csv

rot = here(in_dir, in_rot) %>% 
  read.csv
```

ensure factors are flagged as such
```{r}
facts = 
  c('YEAR', 
    'ROTATION', 
    'REP',
    'CYCLE',
    'CYCLE_YEAR', 
    'PLOT',
    'PREVIOUS', 
    'CURRENT',
    'CYCLE_TEXT')

cycle_text = c(
  `1`='2001\n-\n2004',
  `2`='2005\n-\n2008',
  `3`='2009\n-\n2012',
  `4`='2013\n-\n2016'
)

prev_crop_abbrev = c(
  Canola = 'Ca',
  Corn = 'C',
  Oat = 'O',
  Pea = 'P',
  Soy = 'S',
  Swheat = 'Sw',
  Wwheat = 'Ww'
)

df$CYCLE_TEXT = df$CYCLE %>% 
  as.character %>% 
  cycle_text[.]


is_first_sw = df$YEAR == 2001
df$PREVIOUS[is_first_sw] = NA
df$PREV = df$PREVIOUS %>% 
  as.character %>% 
  prev_crop_abbrev[.] %>% 
  factor(., levels=prev_crop_abbrev)


df[, facts] %<>% lapply(as.factor)
rot[, facts[-c(6:9)]] %<>% lapply(as.factor)




# for plotting
dodge = 0.8
se_hilo = function(x) mean_se(x)[, -c(1)]

# convert yield to Mg/ha
df$YIELD = df$YIELD/1000
rot$YIELD = rot$YIELD/1000
```

# Overall crop yields
```{r}
crop_plots = list()
alt_plots = list()

mod = lmer(YIELD ~ CURRENT*CYCLE_TEXT + (1|YEAR) + (1|REP:PLOT), data=df)
Anova(mod, test.statistic='F')
```

```{r}
emmeans(mod, ~CURRENT|CYCLE_TEXT) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE)
```

Main effects
```{r}
emmm = emmeans(mod, ~CURRENT)
cld(emmm, adjust='fdr', Letters=LETTERS, reverse=TRUE, level=.9)
```
Corn easily yielded the most, followed by winter wheat and oats. Soybean was 
second-worst, well ahead of Canola. 

```{r}
pairs(emmm, adjust='fdr')
```


```{r fig.height=2.5, fig.width=6.5}

within_grp = emmeans(mod, ~CURRENT|CYCLE_TEXT) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE)
colnames(within_grp) %<>% gsub('.group', 'WITHIN', .)

across_grp = emmeans(mod, ~CURRENT) %>% 
  cld(adjust='fdr', Letters=LETTERS, reverse=TRUE)
colnames(across_grp) %<>% gsub('.group', 'ACROSS', .)

grps = merge(within_grp, across_grp[, c('CURRENT', 'ACROSS')], by='CURRENT')
grps[, c('WITHIN', 'ACROSS')] %<>% lapply(trimws)

grps$GROUP = with(grps, paste(WITHIN, ACROSS, sep='\n'))
grps$prettyname = as.character(grps$CURRENT) %>% 
  pretty_crop[.] %>% 
  paste(' (', grps$ACROSS, ')', sep='')
# grps$CURRENT_GRP = with(grps, paste(CURRENT, ' (', ACROSS, ')', sep=''))

pltdf = merge(df, grps[, c('CURRENT', 'prettyname')], by='CURRENT')
pltdf$CYCLE_TEXT %<>% gsub('\n', ' ', .)

grporder = aggregate(YIELD ~ prettyname + CURRENT, data=pltdf, mean)
curorder = grporder[order(grporder$YIELD, decreasing=TRUE), 'CURRENT']
grporder = grporder[order(grporder$YIELD, decreasing=TRUE), 'prettyname']

pltdf$prettyname %<>% factor(levels=grporder)
```


```{r fig.height=2, fig.width=6.5}
pointranges = function(x, r) {
  r = 1-r
  x %<>% na.omit()
  data.frame(y = median(x),
             ymin = quantile(x, r/2),
             ymax = quantile(x, 1-(r/2)))
}

crop_plots[['all']] =
  ggplot(pltdf,
         aes(x=prettyname,
             y=YIELD,
             color=prettyname)) +
  facet_wrap(~CYCLE_TEXT, nrow=1) +
  scale_color_manual(values=crop_colors[curorder] %>% set_names(NULL)) +
  scale_y_continuous(breaks=seq(0, 14, by=2)) +
  # geom_boxplot(shape=21) +
  stat_summary(fun.data=pointranges,
               geom='pointrange',
               fun.args=list(r=.9),
               size=0.1) +
  stat_summary(fun.data=pointranges,
               geom='pointrange',
               fun.args=list(r=0.5),
               linewidth=1.5,
               size=0.5) +
  labs(x=NULL,
       y='Yield [Mg ha<sup>-1</sup>]',
       color=NULL) +
  theme(panel.grid=element_blank(),
        panel.background=element_rect(fill='#f7f7f7', linewidth=0.1),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_line(linewidth=0.1),
        axis.text.x=element_blank(),
        strip.background=element_blank(),
        text=element_text(size=10),
        legend.key=element_blank(),
        axis.title=element_markdown())

crop_plots[['all']]

```

Corn yields more than others, canola the least, and oat and winter wheat are tied for second most, though trail corn by maybe 2000 kg/ha. 

```{r}
"Figure 2 - Crop Yields.jpg" %>% 
  here(out_dir, .) %>% 
  ggsave(crop_plots[['all']], height=2, width=6.5, dpi=300, units='in')
```


# Corn
```{r}
crop = 'Corn'
subdf = subset(df, CURRENT==crop) %>% 
  droplevels

base = subset(subdf, ROTATION=='CS') %>% 
  with(., tapply(.$YIELD, .$YEAR, mean))

subdf$BASE = base[subdf$YEAR]
subdf$DIFF = subdf$YIELD - subdf$BASE

mod = lmer(DIFF ~ ROTATION*CYCLE + (1|YEAR) + (1|REP/PLOT), data=subdf)
Anova(mod, test.statistic='F')
```

Corn yields differently in different rotations. These differences vary across cycles but
not in a substantial way. 

## Contrasts
```{r}
within_grp = emmeans(mod, ~ROTATION|CYCLE) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE) %T>%
  print
```

```{r}
across_grp = emmeans(mod, ~ROTATION) %>% 
  cld(Letters=LETTERS, reverse=TRUE, adjust='fdr') %T>%
  print
```

Corn yielded best in CSSwP and worst in CCSwS. Effects were strongest in Cycles 
1 and 2. 

## Explanations
```{r}
explmod = lmer(DIFF ~ NSPECIES + PREVIOUS + (1|YEAR) + (1|REP/PLOT), 
               data=subset(subdf, YEAR != '2021'))
Anova(explmod, test.statistic='F')
```



```{r}
explmod = lmer(DIFF ~ NSPECIES + (1|YEAR) + (1|REP/PLOT),
               data=subset(subdf, CYCLE != 1))
Anova(explmod, test.statistic='F')
```

### Diversity
```{r}
explmod = lmer(DIFF ~ NSPECIES + (1|YEAR) + (1|REP/PLOT), 
               data=subset(subdf, CYCLE != 1))
Anova(explmod, test.statistic='F')
```

### Previous
```{r}
explmod = lmer(DIFF ~ PREVIOUS + (1|YEAR) + (1|REP/PLOT), 
               data=subset(subdf, CYCLE != 1))
Anova(explmod, test.statistic='F')
```

```{r}
explgrp = emmeans(explmod, ~PREVIOUS) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE) %T>%
  print
```
But corn still does better after pea.

```{r}
emmeans(explmod, ~PREVIOUS) %>% 
  pairs(adjust='fdr')
```


```{r}
summary(explmod)
```
But worse in more diverse rotations after accounting for sequencing effects.

## Plot
Based on those results, we will show two graphs:
1. based on previous crop
2. based on rotation. 
```{r fig.height=2.5, fig.width=6.5}
colnames(within_grp) %<>% 
  gsub('.group', 'WITHIN', .)
colnames(across_grp) %<>% 
  gsub('.group', 'ACROSS', .)

across_grp$ROTATION_GRP = with(across_grp,
                               paste(ROTATION, ' (',
                                     trimws(ACROSS), ')',
                                     sep=''))

grps = merge(within_grp, 
             across_grp[, c('ROTATION', 'ACROSS')], 
             by='ROTATION')
grps[, c('ACROSS', 'WITHIN')] %<>% 
  lapply(trimws)

grps$ROTATION_GRP = with(grps, 
                         paste(ROTATION, ' (', ACROSS, ')', 
                               sep=''))

explgrp$PREVIOUS %<>% as.character
explgrp$PREV_GRP = pretty_crop[explgrp$PREVIOUS] %>% 
  # prev_crop_abbrev[.] %>% 
  paste(' (', 
        trimws(explgrp$.group),
        ')', sep='')

pltdf = 
  subdf %>% 
  merge(grps[, c('ROTATION', 'CYCLE', 'ROTATION_GRP')], 
        by=c('ROTATION', 'CYCLE')) %>% 
  merge(explgrp[, c('PREVIOUS', 'PREV_GRP')], by='PREVIOUS')
lvls = unique(pltdf$ROTATION_GRP) %>% 
  set_names(., strsplit(., ' ') %>% sapply('[', 1)) %>% 
  .[names(rot_colors)]
pltdf$ROTATION_GRP %<>% factor(levels=lvls)

prvgrp_order = order(explgrp$emmean, decreasing=TRUE)
pltdf$PREV_GRP %<>% factor(levels=explgrp$PREV_GRP[prvgrp_order])

rotgrp_order = order(across_grp$emmean, decreasing=TRUE)
pltdf$ROTATION_GRP %<>% factor(levels=across_grp$ROTATION_GRP[rotgrp_order])

# shape_order = across_grp %>% 
#   as.data.frame %>% 
#   set_rownames(.$ROTATION) %>% 
#   .[names(lvls), 'emmean'] %>%
#   order(decreasing=TRUE)

shape_order = rot_shapes[across_grp[, 'ROTATION']] %>% 
  set_names(NULL)

color_order = crop_colors[explgrp[, 'PREVIOUS']] %>% 
  set_names(NULL)
  
```


```{r fig.height=2.5, fig.width=5}
crop_plots[[crop]] = 
  ggplot(pltdf,
         aes(x=as.numeric(CYCLE),
             y=DIFF,
             shape=ROTATION_GRP,
             color=PREV_GRP)) +
  scale_x_continuous(breaks=1:4,
                     labels=cycle_text) +
  # scale_color_brewer(palette='Dark2') +
  scale_color_manual(values=color_order) +
  scale_shape_manual(values=shape_order) +
  geom_hline(aes(yintercept=0),
             color='darkgray',
             size=0.3) +
  stat_summary(fun.data=mean_se,
               geom='pointrange',
               stroke=0.2,
               size=0.5) +
  stat_summary(fun=mean,
               geom='path',
               linewidth=0.2, 
               linetype=3,
               show.legend=FALSE) +
  labs(title=crop, 
       x=NULL,
       y='Relative Corn<br>Yield [Mg ha<sup>-1</sup>]',
       shape='Rotation',
       color='Preceding Crop') +
  guides(color=guide_legend(order=2),
         shape=guide_legend(order=1)) +
  theme(panel.background=element_blank(),
        legend.key=element_blank(),
        strip.background=element_blank(),
        text=element_text(size=10),
        legend.box='horizontal',
        axis.title.y=element_markdown(),
        axis.ticks=element_line(linewidth=0.1))
crop_plots[[crop]]
```

We don't see consistent evidence of corn overyielding in diverse rotations except possibly in CSSwS and CSSwP. This benefit is immediate and variable, but without
an obvious trend. 

# Soybean
```{r}
crop = 'Soy'
subdf = subset(df, CURRENT==crop) %>% 
  droplevels

base = subset(subdf, ROTATION=='CS') %>% 
  with(., tapply(.$YIELD, .$YEAR, mean))

subdf$BASE = base[subdf$YEAR]
subdf$DIFF = subdf$YIELD - subdf$BASE

mod = lmer(DIFF ~ ROTATION*CYCLE + (1|YEAR) + (1|REP:PLOT), data=subdf)
Anova(mod, test.statistic='F')
```

Corn yields differently in different rotations, which varries across year. 

## Contrasts
```{r}
within_grp = emmeans(mod, ~ROTATION|CYCLE) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE) %T>%
  print
```

```{r}
across_grp = emmeans(mod, ~ROTATION) %>% 
  cld(Letters=LETTERS, reverse=TRUE, adjust='fdr') %T>%
  print
```

Soybean yielded better in all 4-year rotations except CCSSw. This effect 
grew as the system matured, so that even CCSSw was a better soybean rotation
than CS. 

## Explanations

```{r}
explmod = lmer(DIFF ~ NSPECIES + PREVIOUS + (1|YEAR) + (1|REP:PLOT), 
               data=subset(subdf, YEAR != 2001))
Anova(explmod, test.statistic='F')
```


```{r}
summary(explmod)
```


### Previous Crop

```{r}
emmeans(explmod, ~PREVIOUS) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE)
```

Soybean does best after spring wheat. 
```{r}
emmeans(explmod, ~PREVIOUS) %>% 
  pairs(adjust='fdr')
```


```{r}
summary(explmod)
```

## Plot
Will plot to show the diversity effect.

```{r fig.height=2.5, fig.width=6.5}
colnames(within_grp) %<>% gsub('.group', 'WITHIN', .)
colnames(across_grp) %<>% gsub('.group', 'ACROSS', .)

grps = merge(within_grp, across_grp[, c('ROTATION', 'ACROSS')], by='ROTATION')
grps[, c('ACROSS', 'WITHIN')] %<>% lapply(trimws)

grps$ROTATION_GRP = with(grps, paste(ROTATION, ' (', ACROSS, ')', sep=''))

pltdf = subdf %>% 
  merge(grps[, c('ROTATION', 'CYCLE', 'WITHIN', 'ROTATION_GRP')], 
        by=c('ROTATION', 'CYCLE'))
lvls = unique(pltdf$ROTATION_GRP) %>% 
  set_names(., strsplit(., ' ') %>% sapply('[', 1)) %>% 
  .[names(rot_colors)]
pltdf$ROTATION_GRP %<>% factor(levels=lvls)


across_grp$ROTATION_GRP = 
  with(across_grp,
       paste(ROTATION, ' (', 
             trimws(ACROSS),
             ')', sep=''))
rotgrp_order = order(across_grp$emmean, decreasing=TRUE)
pltdf$ROTATION_GRP %<>% factor(levels=across_grp$ROTATION_GRP[rotgrp_order])

shape_order = across_grp %>% 
  as.data.frame %>% 
  set_rownames(.$ROTATION) %>% 
  .[names(lvls), 'emmean'] %>%
  order(decreasing=TRUE)

shape_order = rot_shapes[across_grp[, 'ROTATION']] %>% 
  set_names(NULL)

```


```{r fig.height=2.5, fig.width=5}
crop_plots[[crop]] = 
  ggplot(pltdf,
         aes(x=as.numeric(CYCLE),
             y=DIFF,
             color=factor(NSPECIES),
             shape=ROTATION_GRP)) +
  scale_x_continuous(breaks=1:4,
                     labels=cycle_text) +
  scale_shape_manual(values=shape_order) + #rot_shapes[shape_order]) +
  # scale_color_manual(values=c('#fb580c',
  #                             '#e6c000',
  #                             '#4bac34')) +
  scale_color_manual(values=pseudo_palette_3) +
  geom_hline(aes(yintercept=0),
             color='darkgray',
             size=0.3) +
  stat_summary(fun.data=mean_se,
               geom='pointrange',
               stroke=0.2,
               size=0.5) +
  stat_summary(fun=mean,
               geom='path',
               linewidth=0.2, 
               linetype=3,
               show.legend=FALSE) +
  labs(title=crop, 
       x=NULL,
       y='Relative Soybean<br>Yield [Mg ha<sup>-1</sup>]',
       color="Rotation Species",
       shape='Rotation') +
  guides(color=guide_legend(order=2),
         shape=guide_legend(order=1)) +
  theme(panel.background=element_blank(),
        legend.key=element_blank(),
        text=element_text(size=10),
        legend.box='horizontal',
        legend.title=element_markdown(hjust=0),
        axis.title.y=element_markdown(),
        axis.ticks=element_line(linewidth=0.1))
crop_plots[[crop]]
```

Clear evidence of overyielding in diverse rotations that develops over time. 


# Crop-Level Discussion
We see evidence of overyielding for soybean after spring wheat and in diverse rotations. We don't see evidence for overyielding in corn, except following pea, and corn does more poorly in diverse rotations. 


# Rotation: Overyielding?

What is expected rotation yield? = sum(E(yield|crop)) for all crops in a rotation. I can calculate this within each year-rep. 

1. Calculate expected within block based on crop species
2. Calculate observed within block (Fig 1a)
3. Calculate difference

## Calculate expected rotation yield in each rep-year
```{r}
rotations = list(
  CCaWwS = c('Corn', 'Can', 'Wwheat', 'Soy'),
  CCSSw = c('Corn', 'Corn', 'Soy', 'Swheat'),
  COWwS = c('Corn', 'Oat', 'Wwheat', 'Soy'),
  CPWwS = c('Corn', 'Pea', 'Wwheat', 'Soy'),
  CS = c('Corn', 'Soy'),
  CSSwP = c('Corn', 'Soy', 'Swheat', 'Pea'),
  CSSwS = c('Corn', 'Soy', 'Swheat', 'Soy')
)

get_expected = function(x) {
  crop_yields = tapply(x$YIELD, x$CURRENT, mean)
  
  sapply(rotations, function(x) {
    mean(crop_yields[x])
  })
}

eyield = split(df, df[, c('YEAR', 'REP')]) %>% 
  sapply(get_expected) %>% 
  data.frame(ROTATION = rownames(.),
             .) %>% 
  pivot_longer(cols=where('is.numeric'), 
               values_to='EYIELD',
               names_to='YEAR') %>% 
  as.data.frame
eyield[, c('YEAR', 'REP')] = strsplit(eyield$YEAR, '\\.') %>% 
  do.call(rbind, .)
eyield$YEAR %<>% gsub('X', '', .)
```

## Combine with observed rotation yield
```{r}
mergers = c('ROTATION', 'YEAR', 'REP')
for (i in mergers) {
  eyield[, i] %<>% factor(levels=levels(rot[, i]))
}

eyield %<>% merge(rot[, c(mergers, 'CYCLE', 'CYCLE_YEAR', 'YIELD')], 
                  by=mergers, 
                  all.x=TRUE)
eyield$CYCLE_TEXT = eyield$CYCLE %>% 
  as.character %>% 
  cycle_text[.]
eyield$DIFF = with(eyield, YIELD - EYIELD)
```

## Test
```{r}
mod = na.omit(eyield) %>% 
  lmer(DIFF ~ ROTATION*CYCLE + (1|YEAR) + (1|REP/CYCLE_YEAR),
       data=.)
Anova(mod, test.statistic='F')
```


```{r}
within_grp = emmeans(mod, ~ROTATION|CYCLE) %>% 
  cld(adjust='fdr', Letters=letters, reverse=TRUE) %T>%
  print
```

```{r}
across_grp = emmeans(mod, ~ROTATION) %>% 
  cld(Letters=LETTERS, reverse=TRUE, adjust='fdr') %T>%
  print
```

CSSwP showed overyielding, probalby due to the pea-corn benefits. We see weak overyielding in COWwS, CSSwS, and CCaWwS. We see underyielding in CS, CPWwS, and especially CCSSw. Where CCaWwS was successful it generaly was neutral. Effects emerged immediately,
suggesting they're corn-driven. 

```{r}
aggregate(DIFF ~ ROTATION, data=na.omit(eyield), mean)
```
```{r}
aggregate(DIFF ~ ROTATION, data=na.omit(eyield), function(x) sd(x)/sqrt(length(x)))
```

```{r}
aggregate(DIFF ~ ROTATION, data=na.omit(eyield), function(x) t.test(x)$p.value)
```




```{r}
summary(mod)
```


## Explanations
```{r}
subdf = eyield %>% #subset(eyield, CYCLE != 1) %>% 
  merge(unique(df[, c('ROTATION', 'NSPECIES')]), by='ROTATION')

explmod = lmer(DIFF ~ NSPECIES*CYCLE  + (1|YEAR) + (1|REP:CYCLE_YEAR),
               data=subdf)
Anova(explmod, test.statistic='F')
```

```{r}
summary(explmod)
```
 

## Plot
```{r}
colnames(within_grp) %<>% gsub('.group', 'WITHIN', .)
colnames(across_grp) %<>% gsub('.group', 'ACROSS', .)

grps = merge(within_grp, across_grp[, c('ROTATION', 'ACROSS')], by='ROTATION')
grps[, c('ACROSS', 'WITHIN')] %<>% lapply(trimws)

grps$GROUP = with(grps, paste(WITHIN, ACROSS, sep='\n'))
grps$ROTATION_GRP = with(grps, paste(ROTATION, ' (', ACROSS, ')', sep=''))

pltdf = subdf %>%
  merge(grps[, c('ROTATION', 'ROTATION_GRP')], by='ROTATION')

lvls = unique(pltdf$ROTATION_GRP) %>% 
  set_names(., strsplit(., ' ') %>% sapply('[', 1)) %>% 
  .[names(rot_colors)]

pltdf$ROTATION_GRP %<>% factor(levels=lvls)

across_grp$ROTATION_GRP = 
  with(across_grp,
       paste(ROTATION, ' (', 
             trimws(ACROSS),
             ')', sep=''))
rotgrp_order = order(across_grp$emmean, decreasing=TRUE)
pltdf$ROTATION_GRP %<>% factor(levels=across_grp$ROTATION_GRP[rotgrp_order])

shape_order = across_grp %>% 
  as.data.frame %>% 
  set_rownames(.$ROTATION) %>% 
  .[names(lvls), 'emmean'] %>%
  order(decreasing=TRUE)
```


```{r}
crop_plots[['overyield']] = 
  ggplot(pltdf,
         aes(x=as.numeric(CYCLE),
             y=DIFF,
             shape=ROTATION_GRP,
             color=as.factor(NSPECIES))) +
  scale_x_continuous(breaks=1:4,
                     labels=cycle_text) +  
  # scale_color_manual(values=c('#fb580c',
  #                             '#e6c000',
  #                             '#4bac34')) +
  scale_color_manual(values=pseudo_palette_3) +
  scale_shape_manual(values=c(16, 17, 15, 4, 7, 8, 11)[shape_order]) +
  geom_hline(aes(yintercept=0),
             color='darkgray',
             size=0.3) +
  stat_summary(fun.data=mean_se,
               geom='pointrange',
               stroke=0.2,
               size=0.5) +
  stat_summary(fun=mean,
               geom='path',
               linewidth=0.3, 
               linetype=3,
               show.legend=FALSE) +
  labs(x=NULL,
       y='Rotation Yield<br>Deviance [Mg ha<sup>-1</sup>]',
       shape='Rotation',
       color="Rotation Species") +
  guides(shape=guide_legend(order=1),
         color=guide_legend(order=2)) +
  theme(panel.background=element_blank(),
        legend.key=element_blank(),
        strip.background=element_blank(),
        text=element_text(size=10),
        legend.title=element_markdown(hjust=0),
        legend.box='horizontal',
        axis.ticks=element_line(linewidth=0.1),
        axis.title.y=element_markdown())#(orientation='left-rotated'))
crop_plots[['overyield']]
```

CS underyields relative to more diverse rotations. CCSSw definitely underyields.



# Export
```{r fig.height=4.5, fig.width=4.5}
# colors from https://davidmathlogic.com/colorblind/

p1 = crop_plots$all +
  labs(tag='a)') +
  # guides(fill=guide_legend(label.position='bottom',
                           # nrow=1)) +
  theme(text=element_text(size=8),
        strip.text=element_blank(),
        # legend.margin=margin(b=10),
        legend.key.height=unit(0.18, 'in'))

p2 = crop_plots$Corn +
  labs(tag='a)',
       title=NULL) +
  # scale_color_manual(values=c('#EE6677', '#228833', '#CCBB44', '#4477AA')) +
  guides(color=guide_legend(ncol=2, order=1),
         shape=guide_legend(ncol=2,
                            order=2,
                            # override.aes=list(color='#555555')
                            )) +
  theme(text=element_text(size=8),
        axis.text.x=element_blank(),
        # axis.ticks.x=element_blank(),
        legend.key.height=unit(0.12, 'in'),
        legend.box='vertical',
        legend.margin=margin(t=-0.25, unit='cm'))

p3 = crop_plots$Soy +
  labs(tag='b)',
       title=NULL) +
  # scale_color_manual(values=c('#E8601C',
  #                             '#4EB265',
  #                             '#1965B0')) +
  guides(color=guide_legend(ncol=3, order=1),
         shape=guide_legend(ncol=2,
                            order=2,
                            # override.aes=list(color='#555555')
                            )) +
  theme(text=element_text(size=8),
        legend.key.height=unit(0.12, 'in'),
        axis.text.x=element_blank(),
        # axis.ticks.x=element_blank(),
        legend.box='vertical',
        legend.margin=margin(t=-0.25, unit='cm'))

p4 = crop_plots$overyield +
  labs(tag='c)') +
  # scale_color_manual(values=c('#1965B0',
  #                             '#4EB265',
  #                             '#E8601C')) +
  guides(color=guide_legend(ncol=3, order=1),
         shape=guide_legend(ncol=2,
                            order=2,
                            # override.aes=list(color='#555555')
                            )) +
  theme(text=element_text(size=8),
        legend.key.height=unit(0.12, 'in'),
        legend.box='vertical',
        legend.margin=margin(t=-0.25, unit='cm'))
```


```{r fig.height=4.5, fig.width=4.5}
grid.arrange(grobs=list(p2, p3, p4), nrow=3)
```

```{r}
here('Results', 'Figure 4 - Overyielding.jpg') %>% 
  jpeg(width=4, height=4.5, units='in', res=300)
grid.arrange(grobs=list(p2, p3, p4), nrow=3)
dev.off()
```